---
title: "CPU Visualization"
author: "Robert P."
date: "2026-01-03"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Introduction

Comments on code segments will be given *AFTER* the code segment to keep it readable and understandable.

Ex:

```{r}
print("Hello World!")
# This code prints "Hello World!"
```

This document examines the speeds and core count of various CPUs from multiple years to identify any trends between the age of a CPU and its speed and core count.

## Data Import

This section imports our data from the GitHub repository using the raw URL:

```{r}
# Import CSV from GitHub
url <- "https://raw.githubusercontent.com/prunekaryote/CPU-Data-Analysis-Project/refs/heads/main/cpudata.csv"
data <- read.csv(url)
```

## Data Analysis
```{r}
year_avgGHz <- aggregate(GHz ~ Year, data = data, FUN = mean)
year_avgGHz$increase_GHz <- c(NA, diff(year_avgGHz$GHz))

```

Here we calculate the year by year increase in CPU speeds.
```{r}
year_avgCore <- aggregate(Cores ~ Year, data = data, FUN = mean)
year_avgCore$increase_Cores <- c(NA, diff(year_avgCore$Cores))
```
Here the average amount of cores by year is calculated.
<br><br>
We can generate the tables for our data and set them as p1, p2, and p3 using this code:

```{r}
library(tidyverse)
library(patchwork)
library(ggrepel)

p1 <- ggplot(data, aes(Year, GHz)) +
  geom_jitter() +
  geom_text_repel(aes(label = Name)) +
  labs(x = "Year", y = "CPU Speed (GHz)") +
  theme_grey()

# This graph shows each CPU and it's speed by year.

p2 <- ggplot(year_avgGHz, aes(Year, increase_GHz)) +
  geom_col() +
   theme_grey() +
    labs(x = "Year", y = "GHz Increase", 
    title = "Average CPU Speed Increase per Year")

# This graph shows each CPU and it's increase in speed by year.

p3 <- ggplot(year_avgCore, aes(Year, increase_Cores)) +
               geom_col() +
                theme_grey() +
               labs(x = "Year", y = "Core(s) Increase",
                    title = "Average Core Count Increase per Year")

# This graph shows each CPU and it's increase in cores by year.

```
```{r, echo=FALSE}
p1
p2
p3
```



## Conclusion

From the data we can prove a clear trend between the year of CPU release with core and speed increase. This trend shows a standardization of multi-core CPUs after the early 2000s and improvements in processor speed over time.